---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo = F}
library("knitr")
opts_chunk$set(echo = T, results = 'hide')
```

## Loading and preprocessing the data
```{r Read file, echo = T}
activity_data = read.csv("activity.csv")
activity_data$date = as.Date(activity_data$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r}
total_steps = aggregate(steps~date, activity_data, FUN = sum, na.rm = T)
with(total_steps, hist(steps))

options("scipen" = 10)
mean = mean(total_steps$steps, na.rm = T)
median = median(total_steps$steps, na.rm = T)
```

mean =  `r mean`

median =  `r median`

## What is the average daily activity pattern?
```{r, echo = T, results = T}
complete = subset(activity_data, complete.cases(activity_data)==T)
plot(type = 'l', tapply(complete$steps, complete$interval, mean), main = "Average steps taken of each time interval", 
     xlab = "Time interval", ylab = "Averge steps", xlim = c(0, 290))
```

```{r}
minute = names(which.max(tapply(complete$steps, complete$interval, mean)))
index = which.max(unname(tapply(complete$steps, complete$interval, mean)))
```
The `r index`th (`r minute`) interval contains the maximum number of steps, averaged across all the days.


## Imputing missing values
```{r, echo = T}
NA_row = unname(table(complete.cases(activity_data))[1])
```
There are `r NA_row` rows with NAs.

```{r, results = T}
#First find days which measurment never took place.
missing_day = tapply(activity_data$steps, activity_data$interval, function(x)which(is.na(x)))
missing_days = unique(activity_data$date)[missing_day[[1]]]
missing_days

#Find averge for each interval (over entire study period)
average_4_intervals = unname(tapply(activity_data$steps, activity_data$interval, mean, na.rm = T))

#impute missing stpes with average step of that particular interval 
for (day in missing_days){
        activity_data$steps[which(activity_data$date == day)] = average_4_intervals
}

any(is.na(activity_data$steps))

mean = mean(tapply(activity_data$steps, activity_data$date, sum), na.rm = T)
median = median(tapply(activity_data$steps, activity_data$date, sum), na.rm = T)

no_na = aggregate(steps~date, activity_data, FUN = sum, na.rm = T)
with(no_na, hist(steps, main = 'Histogram of step with imputed data'))
```

mean = `r mean`

median = `r median`

The means stays the same, the meandin is slightly higher.


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo = T, results=T}
activity_data$`day of the week` <- ifelse(weekdays(activity_data$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
activity_data$`day of the week` = as.factor(activity_data$`day of the week`)

library(lattice)
weekday = data.frame(Steps = with(activity_data[activity_data$`day of the week`=='Weekday',], unname(tapply(steps, interval, mean))), Interval = unique(activity_data$interval), day = "Weekday")
weekend = data.frame(Steps = with(activity_data[activity_data$`day of the week`=='Weekend',], unname(tapply(steps, interval, mean))), Interval = unique(activity_data$interval), day = "Weekend")
combined = rbind(weekday, weekend)
xyplot(Steps~Interval | day, data = combined, type = "l", layout = c(1, 2)
)
```