---
title: "Activity monitoring data"
author: "YIu Chung Wong"
date: "23/08/2017"
output: html_document
---

```{r setoptions, echo = F}
library("knitr")
opts_chunk$set(echo = F, results = 'hide')
```



##Download data
```{r Download activity monitoring data}
URL = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download_zip = function(URL,  zip_name, output_file_name){
        if(!file.exists(output_file_name)) { #Check wether the folder containing the data sets exists in the working directory.
                if(!file.exists(zip_name)){ # If not then see whether the zip file exist. #If not then download it.
                        switch(Sys.info()[['sysname']],
                               Windows = {download.file(URL, 
                                                        zip_name, method = "wininet", mode = "wb")},
                               Linux  = {download.file(URL, 
                                                       zip_name, method="curl")},
                               Darwin = {download.file(URL, 
                                                       zip_name, method="curl")})
                        unzip(normalizePath(zip_name), exdir=getwd()) #And unzip it
                } else if (file.exists(zip_name)) { # If the zip exists then unzip it.
                        unzip(normalizePath(zip_name), exdir=getwd())
                }
        }    
}

download_zip(URL, "activity.zip", "activity.csv")
```

###Read in data
```{r Read file, echo = T}
activity_data = read.csv("activity.csv")
```

###Reformating dates to date format
```{r reformating to date format, echo = T}
activity_data$date = as.Date(activity_data$date, "%Y-%m-%d")
```




##What is mean total number of steps taken per day?

###Draw plot
```{r Plot histogram of number of steps each day, echo = T, results = T}
barplot(tapply(activity_data[complete.cases(activity_data),1], activity_data[complete.cases(activity_data),2], sum), xaxt="n",
        ylab = "Steps", main = "Steps taken each day\n(2012-10-02 to 2012-11-29)")
mtext(side=1, text="Day 1 to 61\n(excluding day: 1,8,32,35,40,41,45,61)", line=1)
```

###Calculate and report the mean and median total number of steps taken per day
```{r Calculate and report the mean and median total number of steps taken per day, echo = T, results = T}
mean_median = data.frame(Date = unique(activity_data[complete.cases(activity_data),2]), Mean = tapply(activity_data[complete.cases(activity_data),1], activity_data[complete.cases(activity_data),2], mean), Median = tapply(activity_data[complete.cases(activity_data),1], activity_data[complete.cases(activity_data),2], median), row.names = 1:53)
print(mean_median, include.rownames = FALSE)
```




##What is the average daily activity pattern?

###Time series plot 
```{r, echo = T, results = T}
plot(type = 'l', tapply(activity_data[complete.cases(activity_data),1], activity_data[complete.cases(activity_data),3], mean), main = "Average steps taken of each time interval", 
     xlab = "Time interval", ylab = "Averge steps", xlim = c(0, 290))
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo = T}
max_step = which.max(unname(tapply(activity_data[complete.cases(activity_data),1], activity_data[complete.cases(activity_data),3], mean)))
```
The `r max_step`th interval contains the maximum number of steps, averaged across all the days.



##Imputing missing values

###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo = T}
NA_row = unname(table(complete.cases(activity_data))[1])
```
There are `r NA_row` rows with NAs.


###Devise a strategy for filling in all of the missing values in the dataset.
```{r, echo = T, results = T}
#First find days which measurment never took place.
missing_day = tapply(activity_data$steps, activity_data$interval, function(x)which(is.na(x)))
missing_days = unique(activity_data$date)[missing_day[[1]]]

#Find averge for each interval (over entire study period)
average_4_intervals = unname(tapply(activity_data[complete.cases(activity_data),1], activity_data[complete.cases(activity_data),3], mean))

#impute missing stpes with average step of that particular interval 
for (day in missing_days){
        activity_data$steps[which(activity_data$date == day)] = average_4_intervals
}

print(head(activity_data))
```


##Histogram of the total number of steps taken each day
```{r, echo = T, results = T}
barplot(tapply(activity_data[,1], activity_data[,2], sum), xaxt="n",
        ylab = "Steps", main = "Steps taken each day\n(2012-10-02 to 2012-11-29)")
mtext(side=1, text="Day 1 to 61", line=1)

mean_median = data.frame(Date = unique(activity_data[,2]), Mean = tapply(activity_data[,1], activity_data[,2], mean), Median = tapply(activity_data[,1], activity_data[,2], median), row.names = 1:61)
print(mean_median, include.rownames = FALSE)
```
For non missing days, the value stays the same. For days without measurments, their median is now the median of all the averaged steps (over the study period) within each interval.
They also happen to have the same mean because they have exactly the same data. 




##Are there differences in activity patterns between weekdays and weekends?
```{r, echo = T, results=T}
activity_data$`day of the week` <- ifelse(weekdays(activity_data$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
activity_data$`day of the week` = as.factor(activity_data$`day of the week`)

library(lattice)
weekday = data.frame(Steps = with(activity_data[activity_data$`day of the week`=='Weekday',], unname(tapply(steps, interval, mean))), Interval = unique(activity_data$interval), day = "Weekday")
weekend = data.frame(Steps = with(activity_data[activity_data$`day of the week`=='Weekend',], unname(tapply(steps, interval, mean))), Interval = unique(activity_data$interval), day = "Weekend")
combined = rbind(weekday, weekend)
xyplot(Steps~Interval | day, data = combined, type = "l", layout = c(1, 2)
)
```
